The Social Conditioning of Fear

### 2023-12-06 scripts/manuscript.qmd
transgender

This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue. With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among the US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education. We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walkign alone at night" and our other variables of interest.

@@

First, we examined how fear is experienced across race and gender. As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents who identified as Male. Particularly, respondents that were Transgender expressed greater fear, but this may be indicative of sample representation. For example, only one respondent identified as both Hispanic and Transgender, and since they responded affirmative to the question, this finding is heavily skewed. When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest. This finding is consistent with the hypothesis that the most privileged social group in terms of race has the lowest level of fear.

>>>>

This is an exploratory project for a dissertation chapter idea that one of the co-authors wants to pursue. With descriptive statistics, we are interested in investigating the inequalities in terms of fear of walking alone at night among the US adult population and mapping out the patterns of disparities along the axes of race, gender, sexual orientation, citizenship status, and education. We use the 2022 data from the General Social Survey (GSS) and the data science tools from EDLD 651 class to delineate the data distribution and basic correlations between "fear of walking alone at night" and our other variables of interest.

@@

First, we examined how fear is experienced across race and gender. As @tbl-FearRaceSex shows, Transgender and Female respondents had higher mean fear rate than respondents who identified as Male. Particularly, respondents that were Transgender expressed greater fear, but this may be indicative of sample representation. For example, only one respondent identified as both Hispanic and Transgender, and since they responded affirmative to the question, this finding is heavily skewed. However, it's important to note that Transgender individuals have been subject to more frequent and more severe forms of sexual violence and thus, experiencing more fear related to gender-based crime and engaging in more gendered tactics to protect themselves [@schilt2009].

When looking by race, whites were the group with the lowest mean fear and Asians were the group with the highest. This finding is consistent with the hypothesis that the most privileged social group in terms of race has the lowest level of fear.
